Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2022-07-14 ba_resubmit/response_to_reviewers.tex

\documentclass[letterpaper, parskip]{scrartcl}

\usepackage{amsmath,amssymb,bm}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[margin=1.4in]{geometry}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{tcolorbox}            % provides shaded box for reviewer comments
\usepackage{enumitem}             % use more compact lists
%\usepackage[parfill]{parskip}     % use skips between paragraphs and no indents
\usepackage{bibentry}             % provides \nobibliography command
\usepackage{xspace}

% Make itemize and enumerate environments more compact
\setlist[itemize]{noitemsep, topsep=0pt, parsep=0pt}
\setlist[enumerate]{noitemsep, topsep=0pt, parsep=0pt}

% -------- Define point raised command --------- %

\definecolor{darkpurple}{HTML}{500050}
\tcbuselibrary{breakable}       % allow tcolorbox to break across pages
\tcbuselibrary{skins}
\tcbset{%
	enhanced,%
	frame hidden,%
	sharp corners,%
	breakable,%
	parbox=false,%
	colback=lightgray!30,%
	enlarge left by=6pt,%
	width=\linewidth-6pt,%
	colback=gray!0,
	left=4pt,%
	right=0pt,%
	top=0pt,%
	bottom=0pt,%
	borderline west={2pt}{0pt}{lightgray},%
	%coltext=darkpurple,%
	before skip=14pt plus 2pt,%
	after skip=14pt plus 2pt}

\newcommand{\pointRaised}[1]{%
	\begin{tcolorbox}
		\itshape #1
	\end{tcolorbox}
}

\newcounter{responsectr}[section]     % counter for response resets at each section
\newcommand{\reply}[2]{%
	\refstepcounter{responsectr}%
	\textbf{#1.\theresponsectr:} #2
}

\newcommand{\todo}{\textcolor{red}{[TODO]}\xspace}

\newcommand{\ncomment}[1]{{\color{teal} NM: #1}}

\begin{document}

	% ****************** TITLE ****************************************

	\title{Response to Reviewer Comments}

	\maketitle
	We thank the Editor and reviewers for their encouraging assessment and constructive 
	feedback.

	\section*{Editor}

	Dear Dr Kundinger,

	Thank you for submitting your paper "Efficient and Scalable Bipartite Matching through Fast Beta Linkage (fabl)" for possible publication in Bayesian Analysis.
	It has now been carefully reviewed and my decision is: Major revision required.

	These are the comments I received from the Editor who has handled your paper:

	The paper has been reviewed by a referee and an associate editor and I have read the manuscript independently, before looking at the reports.  When I confront my notes and overall impression with the reviewers’ comments, I find myself in substantial agreement with the reviewers.

	As indicated by both reviewers, the work expands on previous work by Sadinle, introducing computational advances that are achieved by relaxing some requirements in the original formulation of the problem.  The close connections with the previous work and the extent of the novel contributions are not adequately explained.  This makes it hard for the reader to understand what is new and why the new contributions are valuable.  A careful editing of the presentation is needed to address this shortcoming.

	Regarding the presentation, the AE points out a number of problems with notation, typos, and various inconsistencies.  Independently, I found many similar problems.  The intersection between what I found and what the AE found is non-empty and so is the symmetric difference.  Also, the AE found some issues that I did not find and vice versa.  So, I am almost certain that an additional review would uncover more problems.  Now, the results presented in the paper seem plausible as do the broad strokes of the derivations, but I must admit that I was not able to follow all the details. This, in large part, was because of the issues that I just mentioned, and I believe that most readers would find it difficult to follow the developments, as these issues are exceedingly distracting.

	There are problems with the notation, which in places is not defined, in others is used before being defined, and in others yet is used inconsistently.  For example, I_obs() is never defined, n_12(Z) first appears on p. 5, but is not defined (in passing) until the bottom of p. 7, and it seems to become D on p. 8.  Throughout the manuscript, n_1 and n_2 are used interchangeably with n_A and n_B, sometimes in the same section, as it happens, for example, in Section 3.1.  The loss function on p. 9 is out of Sadinle’s paper and it inherits the typo therein (\theta_11, if Z_j, \hat{Z_j}, …).  In the statement of Lemma 1 on p.13, B comes out of nowhere.  (Thinking I had missed something, I went back to the previous pages only to find out that B would be then defined in the proof.)  Formatting of all displayed math must be checked and fixed, especially as far as punctuation is concerned, as commas seem to be missing in multiple places.  For example, in Equation (1), there should be a comma after 0 and after 1, and this is by no means an isolated occurrence.

	Confusion can also arise from lack of clarity in the exposition.  Take comment 4 from the AE, for example.  Whose marriage certificates and whose birth records are involved?  What records are in X_1 and what records are in X_2?  The reader should not be expected to go back to Newcombe’s paper to clarify the issue.

	This is not an exhaustive list, but only a set of examples meant to point out what makes the paper hard to read and how it can be improved.  Many more specific points are raised in the AE’s report and here are a couple of other minor issues I noticed:

	p. 12, second paragraph:  insert “is” after Section 3.1

	p. 15:  the first sentence of the last full paragraph is either missing a verb (“we observe”?) or it includes an extra “that.”

	As I said, I do not believe that we, as reviewers, were able to uncover all the problems, and I strongly encourage the authors to do their part, as they should, to improve the presentation and eliminate all typos and inconsistencies.

	The AE finds the simulation studies to be incomplete and gives detailed suggestions on how those should be improved.  The AE also gives important suggestions about other aspects of the manuscript that must be carefully considered and addressed.  All of the AE’s comments are right on the mark.

	To summarize, the paper contains an interesting algorithmic contribution that can speed up calculations at the expense of relaxing some of the modeling conditions, without much adverse impact on the resulting inferences.  This aspect should be emphasized in a revised presentation.  The presentation should also make clearer the close connections to the relevant work by Sadinle, and all the issues that I mentioned above should be ironed out.

	The authors should prepare a careful and substantive revision that remains within the editorial limit of 25 pages and that answers the various comments satisfactorily.  The authors should upload the revised manuscript together with a document detailing how they addressed the reviewers’ comments.

%	\reply{E}{%
%		Thank you for the opportunity to submit a revision. 
%		We have provided detailed responses to the reviewers' comments below, 
%		which prompted significant changes in our paper.
%	}

	\clearpage
	\newpage

	\section*{Reviewer 2}
	\setcounter{responsectr}{0}
	This manuscript addresses the bipartite matching problem in data linkage and proposes a new fast
	computation version of “Beta Linkage” to allow such matching to take place for pairs of large databases
	without resorting to ad hoc blocking. It is built upon the Felegi-Sunter model of record linkages, where
	linkages are made within strata of agreement between records across the two databases, and relies
	heavily on advances made by Sadinle (2017).
	Overall, the paper presents a practical implementation of a modified Beta Linkage approach to
	matching. It demonstrates that even though the new method suffers from the limitations of not
	seamlessly enforcing one-to-one matching, it is feasible for large datasets with modest numbers of
	features on which to match. Though the novelty of the presented methodology is primarily in the
	computational approach, I believe it to be a real advance for practical implementation of record linkage
	in modern applications.
	I have no major concerns about the content or accuracy of the presentation. However, substantial
	improvements are warranted to confirm my impression. In addition to the recommendation made by
	Reviewer 1 to include more clarity of the novel contributions of this manuscript, I believe that the
	explanation of the method should be written more clearly, the simulation comparisons should be
	repeated/expanded, and various features of the simplification should be fleshed out a bit more. Finally,
	more practical advice would be welcome. All of these points are more specifically described in the
	itemized list below

		\pointRaised{%
Overall, the notation is, somewhat by necessity, quite extensive. However, I’m not convinced
that it’s entirely consistent. A glossary in the appendix may help the reader navigate the paper
more easily.
	}

		\pointRaised{%
Section 2, line 1: It may help to explicitly define X1 and X2 as being vectors of indices as implied
by equations (1) and (2).
}
		\pointRaised{%
Section 2.2, beta distribution: the n_{12}(.) notation needs to be defined.
}

		\pointRaised{%
Section 3, 4th full paragraph: My understanding of the example of matching birth records and
marriage certificates (presuming monogamy) may not quite be consistent with the algorithm as
presented, as I would think that one certificate (element in X1) would match multiple birth
records (elements in X2) which violates the size ordering. Can the algorithm be easily adjusted
for this reverse sizing?
}
		\pointRaised{%
5. Section 3, full model: I_obs(.) seems to be taken from a different paper dealing with missing
fields, which is not at all discussed in this manuscript. This consideration of missing fields and
the accompanying discussion of missing data mechanism should be included if such data is to be
considered for this manuscript. This comment also applies to the Gamma^{obs} notation in the
following subsection.
}
		\pointRaised{%
Section 3, full model: The Phi notation needs to be defined.
}
		\pointRaised{%
Section 3, full model: The square brackets seem to be misplaced, as the exponent includes
indices of the sum contained in the brackets.
}
		\pointRaised{%
Section 3.1, I found the motivational descriptions in this section to be somewhat confusing, in
that they seem to mix the posterior behavior (e.g., parameters are updated though standard
multinomial-Dirichlet…) and model construction (as a function of pi, Z is a series of successes…).
This subsection might be easier to understand if its purpose was more bluntly described in an
opening sentence or paragraph and streamlined to focus only on the full conditional
distributions for a Gibbs Sampler. In addition, the mixed motivation may have led to errors in
intermediate conclusions. In particular, the derivation of the probability of Gamma_{.j} seems
to be missing a factor of u_{fl}^I(gamma_{ij}^f=l). This is not important to the eventual full
conditional of Z, but does not seem to be correct for the distribution of Gamma as claimed.
}
		\pointRaised{%
Section 3.1, full joint posterior distribution: I was surprised to see that the contribution of the
conditional distribution of Z was included using the summation-style notation rather than the
binomial/multinomial-style multiplicative notation that naturally motivates the full conditional
presented on the first line of page 8.
}
		\pointRaised{%
Section 3.1, first equation on page 8: There is an extra parenthesis in the exponent of (1-pi).
}
		\pointRaised{%
Section 3.1, pmf for Gamma_{.j} and full conditional for Z: The notation in the final equation
may be more clear if you replace i with z_j, as in w_{z_j j}. I believe that “n_1+1” should be “n_1
+ j” in several places.
}
		\pointRaised{%
Section 3.1, last equations on page 8: There seems to be some lack of specificity in describing
these equations as full conditional distributions vs. steps in the Gibbs Sampler algorithm. It
would be helpful to provide a brief justification of integrating out pi.
}
		\pointRaised{%
Section 4.2, SEI procedure. I’m not sure I follow the description of the SEI procedure. My
understanding of SEI is that for patterns with lots of possible pairs, no single pair is likely to
(should) be identified in the posterior distribution. Thus, you take a small subsample of the
records of size S < H_j_p, and store these in R^{SEI} rather than the complete index of pairs in
each R. Can you quantify the computational savings due to this method? Do you have advice for
how small S can be? In your simulations, how does the use of this method or choice of S affect
the results?
}
		\pointRaised{%
Section 4.4, Assumptions and definitions should be included in the statement of this lemma.
}
		\pointRaised{%
Section 5.1 simulations – there is a detailed comparison for one set of simulation settings,
including a binary definition of a “match”. Do these results differ as these settings change?
}
		\pointRaised{%
Section 5.2 and 5.3 simulations – are these a single simulated datasets? Do these same results
hold over multiple simulated datasets, Or was this just a chance result for each setting?
}
		\pointRaised{%
Section 5.3 simulation – RR results are not presented.
}
		\pointRaised{%
	Section 6.1 last sentences. I believe the last sentence should be removed, as it is redundant.
}
		\pointRaised{%
Section 6.2, second sentence. I believe the word “survey” is missing, as in “the initial SURVEY
began…”
}
		\pointRaised{%
Section 6.2. In the NLTCS, are participants added in later cohorts, or may we assume that every
member of the smaller dataset should also be included in the earlier one?
}
		\pointRaised{%
	Figures 5 and 7 are not referenced in the text.
}

	\clearpage
	\newpage

	\section*{Reviewer 3}
	\setcounter{responsectr}{0}

	The paper presents a computational variant of the method introduced by Sadinle (2017). Simulation
	results are presented and the variant is applied on a new, significantly larger data set.
	A significant part of the paper is a reproduction of the paper of Sadinle. The variant is introduced in
	section 4. It is simple but consequential. The simulations and applications show how the variant
	accelerates the computation relative to the original method and reveals the trade-offs.
	The value of the paper resides in that computational complexity continues to be a difficult obstacle in
	the application of Bayesian statistics. The proposals of the paper enable the application of Sadinle’s
	approach to an extent not possible before.
	It is important to present the research for what it is: An extension of the method of Sadinle motivated
	by computational considerations. Here Sadinle supersedes Fellegi-Sunter in the sense while the starting
	point of Sadinle’s research is Fellegi-sunter, the starting point of this paper is Sadinle’s work. This should
	be made clearer, beginning in the abstract and reiterated in the conclusion.

\end{document}

### 2022-09-01 ba_resubmit/kundinger_resubmit.tex
Markdown file for addressing reviewer comments

One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - \bm{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

\begin{align*}
	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\left[ \prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}) \\
	Z_j | \pi &=
	\begin{cases} 
		\frac{1}{n_1}\pi  & z_j \leq n_1; \\
		1-\pi &  z_j  = n_1 + j \\
	\end{cases} \\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}).
\end{align*}

@@

As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_A + 1\)) and failures (when \(z_j = n_A + 1\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$. We have

\begin{align*}
	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - (n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
\end{align*}
Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(D + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.

@@

\begin{align*}
	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	&=
	\begin{cases} 
		w_{ij}  & z_j \leq n_1; \\
		1 &  z_j  = n_1 + j \\
	\end{cases}\\
\end{align*}

@@

With the likelihood in this form, we can derive the full conditional
\begin{align*}
	p(Z_j|\Gamma_{.j}, \Phi, \pi) &\propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\pi) \\
	&\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right) \\	
	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}.
\end{align*}
Thus, we have
$$Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
	\frac{\pi}{n_1}w_{ij}   & z_j \leq n_1; \\
	1-\pi &  z_j  = n_1 + 1. \\
\end{cases}$$
We integrate over \(\pi\) and rearrange terms to produce the final full conditional:

\[p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
\begin{cases} 
	w_{ij}  & i \leq n_1; \\
	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & i  = n_1 + 1. \\
\end{cases}\]

@@

We calculate a Bayes estimate \(\hat{\bm{Z}}\) for the linkage
parameter \(\bm{Z}\) by assigning different positive losses to
different types of errors, and minimizing posterior expected loss. We
adopt the loss function proposed in \cite{sadinle_bayesian_2017}, in which
\(\hat{Z}_j \in \{1, \ldots, n_A + 1, R\}\), with \(R\) representing the
option to leave the matching undetermined by the model. Specifically, we have

\[L(\hat{Z_j}, Z_j)=\begin{cases} 
	0  & \text{if } Z_j = \hat{Z_j}; \\
	\theta_R,  & \text{if } \hat{Z_j} = R; \\
	\theta_{10},  & \text{if } Z_j \leq n_A,\hat{Z_j} = n_A + 1 ; \\
	\theta_{01},  & \text{if } Z_j = n_A + 1,\hat{Z_j} \leq n_A ; \\
	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_A, Z_j \neq \hat{Z_j}. \\
\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
on the linkage status, \(\theta_{10}\) is the loss from a false
non-match, \(\theta_{01}\) is the loss from a false match, and
\(\theta_{11}\) is the loss from the special case of a false match in
which the record has a true match other than the one estimated by the
model.

@@

Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities. These are simple calculations, but become computationally burdensome when working on large linkage tasks.

Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote $H_p^m = \sum_{j=1}^{n_B} \bm{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $p$. The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\). Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes

@@

Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} conceptually straightforward, it becomes computationally burdensome when $n_1$ is large. This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$. To speed up computation, we break this sampling step into two. First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link. Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + 1\). Then,

@@

	\begin{lemma}
		The overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P).$
		\label{lemma:fabl}
	\end{lemma}

@@

	\begin{figure}[t]
		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_A$ and $n_B$. We see near quadratic growth in run-time for BRL, and near linear growth for fabl.}\label{fig:speed1}
		\end{center}
	\end{figure}

@@

	\begin{figure}[h!]
		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_A$, and $n_B$ fixed at 500. We see linear growth in run-time for BRL, and near constant run-time for fabl.}\label{fig:speed2}
		\end{center}
	\end{figure}

@@

	We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review. We note that Bayes estimates falling outside of posterior credible intervals has occurred previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.

	We note that reconciling Bayes estimates that fall outside of posterior credible intervals is a common challenge in record linkage and an avenue of future work.

>>>>

One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12} = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

\begin{align*}
	\mathcal{L}(\bm{Z}, \Phi \mid \bm{\Gamma}) &= \prod_{j=1}^{n_1}  \prod_{i=1}^{n_2}\prod_{f=1}^{F}\prod_{l=1}^{L_f}\left[  m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	\bm{m}_f &\sim \text{Dirichlet}(\alpha_{f0}, \ldots, \alpha_{f L_f}) \\
	\bm{u}_f &\sim \text{Dirichlet}(\beta_{f0}, \ldots, \beta_{f L_f}) \\
	Z_j | \pi &=
	\begin{cases} 
		\frac{1}{n_1}\pi  & z_j \leq n_1; \\
		1-\pi &  z_j  = n_1 + j \\
	\end{cases} \\
	\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}),
\end{align*}
where we have used the condensed notation $\Phi = \{\bf{m}, \bf{u}\}$ to denote the set of parameters governing the probabilities of agreement levels for matching and non-matching record pairs.

@@

Since $\pi$ encodes the rate of matching across the two datasets, the posterior distribution $p(\pi|\Gamma, \bf{Z}, \Phi, \alpha_{\pi}, \beta_{\pi})$ depends only on the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$ (and hyperparameters). Thus we use the abbreviated $p(\pi | \bf{Z}, \alpha_{\pi}, \beta_{\pi})$, and have

%As a function of \(\pi\), \(\bm{Z}\) is sequence of successes (when \(z_j < n_1 + j\)) and failures (when \(z_j = n_1 + j\)), and therefore \(p(\bm{Z}|\pi) = \mathcal{L}(\pi|\bm{Z})\) is determined only by the number of links $n_{12}(\bm{Z}) = \sum_{i=1}^{n_2}I(z_j < n_1 + 1)$ encoded by $\bm{Z}$. We have

\begin{align*}
	p(\pi | \bm{Z}, \alpha_{\pi}, \beta_{\pi}) &\propto p(\bm{Z}|\pi)p(\pi) \\
	&\propto \pi^{n_{12}(\bm{Z})} (1-\pi)^{n_2 - n_{12}(\bm{Z})} \pi^{\alpha_{\pi} -1} (1-\pi)^{\beta_{\pi} -1} \\
	&\propto \pi^{n_{12}(\bm{Z}) + \alpha_{\pi} - 1} (1-\pi)^{n_1 - n_{12}(\bm{Z}) + \beta_{\pi} -1}. \\
\end{align*}
Thus $\pi^{(s+1)}|\bm{Z}^{(s+1),  \alpha_{\pi}, \beta_{\pi}}$ has a $\text{Beta}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})$ distribution.

@@

\begin{align*}
	p(\Gamma_{.j}|Z_j = z_j, \Phi) &\propto \prod_{i=1}^{n_1}\left[\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(Z_j = i)}u_{fl}^{I(Z_j \neq i)}\right]^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}\\
	&\propto \prod_{i=1}^{n_1}\left(\prod_{f=1}^{F}\prod_{l=1}^{L_f} \frac{m_{fl}}{u_{fl}}\right)^{I(z_j = i) I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	&=
	\begin{cases} 
		w_{z_j, j}  & z_j \leq n_1; \\
		1 &  z_j  = n_1 + j \\
	\end{cases}\\
\end{align*}

@@

With the likelihood in this form, we can derive the full conditional
\begin{align*}
	p(Z_j|\Gamma_{.j}, \Phi, \pi) &\propto p(\Gamma_{.j}| Z_j, \Phi) P(Z_j|\pi) \\
	&\propto \left(\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + \bm{1}_{z_j = n_1 + 1}\right)\left(\pi\sum_{i=1}^{n_1}\frac{1}{n_1}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}\right) \\	
	&= \frac{\pi}{n_1}\sum_{i=1}^{n_1}w_{ij}\bm{1}_{z_j = i} + (1-\pi)\bm{1}_{z_j = n_1 + 1}.
\end{align*}
Thus, we have
$$Z_j^{(s+1)} | \Phi, \Gamma, \pi \propto \begin{cases} 
	\frac{\pi}{n_1}w_{z_{j},j}   & z_j \leq n_1; \\
	1-\pi &  z_j  = n_1 + 1. \\
\end{cases}$$
We integrate over \(\pi\) and rearrange terms to produce the final full conditional:

\[p\left(Z_j^{(s+1)}  = i| \Phi, \bm{Z^{(s)}}\right) \propto
\begin{cases} 
	w_{ij}  & i \leq n_1; \\
	n_1 \frac{n_2 - n_{12}(\bm{Z}) + \beta_{\pi}}{n_{12}(\bm{Z}) + \alpha_{\pi}} & i  = n_1 + j. \\
\end{cases}\]

@@

We calculate a Bayes estimate \(\hat{\bm{Z}}\) for the linkage
parameter \(\bm{Z}\) by assigning different positive losses to
different types of errors, and minimizing posterior expected loss. We
adopt the loss function proposed in \cite{sadinle_bayesian_2017}, in which
\(\hat{Z}_j \in \{1, \ldots, n_1 + j, R\}\), with \(R\) representing the
option to leave the matching undetermined by the model. Specifically, we have

\[L(\hat{Z_j}, Z_j)=\begin{cases} 
	0  & \text{if } Z_j = \hat{Z_j}; \\
	\theta_R,  & \text{if } \hat{Z_j} = R; \\
	\theta_{10},  & \text{if } Z_j \leq 1,\hat{Z_j} = n_1 + j ; \\
	\theta_{01},  & \text{if } Z_j = n_1 + j,\hat{Z_j} \leq n_1 ; \\
	\theta_{11},  & \text{if } Z_j, \hat{Z}_j \leq n_1, Z_j \neq \hat{Z_j}. \\
\end{cases}\] Here, \(\theta_R\) is the loss from not making a decision
on the linkage status, \(\theta_{10}\) is the loss from a false
non-match, \(\theta_{01}\) is the loss from a false match, and
\(\theta_{11}\) is the loss from the special case of a false match in
which the record has a true match other than the one estimated by the
model.

@@

Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\bm{Z})\) and \(\beta_{fl}(\bm{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_1 \times n_2\) quantities. These are simple calculations, but become computationally burdensome when working on large linkage tasks.

Instead, we use one-hot encodings of the agreement patterns $\mathcal{P}$ for more efficient calculations. Denote $H_p^m = \sum_{j=1}^{n_2} \bm{1}_{(Z_j, j) \in h_p}$ to be the number of matching record pairs with agreement pattern $p$. The number of non-matching record pairs with agreement pattern $p$ is \(H_p^u = H_p - H_p^m\). Then, if $\bm{\alpha_0}$ and $\bm{\beta_0}$ are vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively, the posterior update becomes

@@

Although sampling \(Z_j\) from a the full conditional provided in Section~\ref{gibbs_sampling} conceptually straightforward, it becomes computationally burdensome when $n_1$ is large. This is because sampling a value from $n_1$ many options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_1$. To speed up computation, we break this sampling step into two. First, we calculate the Fellegi Sunter weight \(w_{p}\) associated with each unique pattern and sample the agreement pattern between \(j\) and its potential link. Second, we sample the record label uniformly among records associated with that agreement pattern for that particular \(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement pattern between \(j\) and its potential match, and say \(h(Z_j) = h_{P+1}\) when \(Z_j = n_1 + j\). Then,

@@

	\begin{lemma}
		The overall computational complexity of \texttt{fabl} is $O(\frac{F}{B} n_1 n_2) + O(n_2 P)$, where $B$ is the number of processors available for parallelization.
		\label{lemma:fabl}
	\end{lemma}

@@

	\begin{figure}[t]
		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/sadinle_speed_plot2} 
			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, for increasing values of both $n_1$ and $n_2$. We see near quadratic growth in run-time for BRL, and near linear growth for fabl.}\label{fig:speed1}
		\end{center}
	\end{figure}

@@

	\begin{figure}[h!]
		\begin{center} \includegraphics[width=0.6\textwidth]{../notes/figures/speed_plot_fixed_nB_slides} 
			\caption{Run-time for BRL and fabl to run 1000 Gibbs iterations, including hashing step for fabl, with increasing $n_1$, and $n_2$ fixed at 500. We see linear growth in run-time for BRL, and near constant run-time for fabl.}\label{fig:speed2}
		\end{center}
	\end{figure}

@@

	We see similar results under \texttt{BRL}, with a Bayes estimate of 181 individuals recorded in both datasets, a posterior 95\% credible interval of (211, 244), and a range of (140, 294) after the partial estimate and clerical review. See `\ref{fig:overlap-plot} for a visual comparison of the Bayes estimate and posterior credible intervals for the two methods. We note that Bayes estimates falling outside of posterior credible intervals has occurred previously in the record linkage literature \citep{sadinle_bayesian_2017, steorts_bayesian_2016}, and remains a topic for future research.
